---
title: "Analysis"
format: html
---
Below are a few graphs which relate to the main graph on the home page.


## Industrialization
```{r}
#| label: setup
#| message: FALSE
#| echo: FALSE
# loading in the libraries
library(tidyverse)
library(leaflet)
library(ggplot2)
library(sf)
library(dplyr)
library(viridis)
library(maps)
library(countrycode)
library(mapproj)
```

```{r}
#| label: data-download
#| message: FALSE
#| echo: FALSE
# reading in the data
industrialization <- read_csv(file = "data/industrialization.csv", show_col_types = FALSE)

# filtering out any bad/unuseful data
industrialization <- industrialization |> 
  filter(Indicator == "Industrialization intensity index") |> 
  select(Emissions, country_code, country_name)

industrialization$country_name[industrialization$country_code == "USA"] <- "United States"
industrialization$country_name[industrialization$country_code == "GBR"] <- "United Kingdom"

#importing in the map
industrialization$country_name <- countrycode(industrialization$country_code, origin = "iso3c", destination = "country.name")

# Load world map data
world_map <- map_data("world")
world_map_regions <- unique(world_map$region)

# Join the map data with the industrialization data
industrialization <- industrialization %>%
  mutate(region = ifelse(country_name %in% world_map_regions, country_name, country_code))
ind_map_data <- left_join(world_map, industrialization, by = "region", relationship =
  "many-to-many")

```

```{r}
#| label: plot-heatmap
#| message: false
#| echo: false
#| warning: false

# dimensions for graph
#| fig-width: 30
#| fig-height: 24

#plot
ind_map_data |> ggplot(aes(x = long, y = lat, group = group, fill = Emissions)) +
  geom_polygon(color = "black", linewidth = 0.2) +
  scale_fill_viridis(name = "Industrialization Intensity Index", 
                     label = scales::comma, 
                     option = "viridis", 
                     direction = -1) +
  labs(
       title = "Industrialization Index by Country (2019)",
       subtitle = "Grey means data not available",
       caption = "Source: Competitive Industrial Performance Index",
       x = NULL,
       y = NULL,) + 
  coord_map("mollweide", 
            xlim = c(-180, 180),
            ylim = c(-90, 90)) + 
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) 


```

The Industrial Intensity Index an index that indicates the performance of various industrial sectors of the economy. In other words, it measures the overall country developedness and production of material goods by country. We can see here that countries such as China, India, Germany, Indonesia, Canada, Russia and America all have fairly high Industrial Index values. Some of these countries have high emission values as shown on the initial graph on the home page.


## Carbon Emissions by Year

```{r}
#| label: data-download-pt-2
#| message: FALSE
#| echo: FALSE

#downloading data
world <- read_csv(file = "data/carbon-emissions.csv", show_col_types = FALSE)

#filtering out just the world data
world <- world |> filter(country_code == "WLD")
```

```{r}
#| label: second-plot
#| message: FALSE
#| echo: FALSE

# dimensions for graph
#| fig-width: 20
#| fig-height: 20

# plot
world |> 
  ggplot(aes(x = year, y = value))  + 
    geom_line() +
    geom_point() +
  scale_y_continuous(labels = scales::comma) +
  scale_x_continuous(breaks = c(1960, 
                                1965, 
                                1970,
                                1975, 
                                1980, 
                                1985, 
                                1990, 
                                1995, 
                                2000, 
                                2005, 
                                2010, 
                                2015, 
                                2020)) + 
    labs(title = "Overall carbon emissions by year",
         subtitle = "We can see rapid growth in the last 60 years",
         x = "Year",
         y = "Kilotons of Carbon Dioxide Emitted",
         caption = "Source: UNFCC") + 
    theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.ticks.x = element_line(color = "black"))
```

In this graph, we see that the amount of carbon emitted every year since 1960 has increased steadily. This can be due to industrialization once again, since the production of carbon emitting items (cars, factories, plants) has been known to incerase every year. In the years 2008 and 1990, there were major economic setbacks which limited the production of goods and as a result, there were lower carbon emissions those years as shown on the graph.

Additionally, here are the yearly Carbon emissions by country.


```{r}
#| label: data-download-pt-3
#| message: FALSE
#| echo: FALSE

#downloading data
countries <- read_csv(file = "data/carbon-emissions.csv", show_col_types = FALSE)

#filtering out just the world data
countries <- countries |> filter(country_code %in% c("CHN", "USA", "RUS", "IDN", "JPN", "KOR"))

```

```{r}
#| label: third-plot
#| message: FALSE
#| echo: FALSE



# plot
countries |> 
  ggplot(aes(x = year, 
             y = value,
             color = country_name,
             group = country_name)) +
    geom_point(size = 0.5) + 
    geom_line() +
  scale_y_continuous(labels = scales::comma,
                     breaks = c(1500000,
                                3000000,
                                4500000,
                                6000000,
                                7500000,
                                9000000)) +
  scale_x_continuous(breaks = c(1960, 
                                1965, 
                                1970,
                                1975, 
                                1980, 
                                1985, 
                                1990, 
                                1995, 
                                2000, 
                                2005, 
                                2010, 
                                2015, 
                                2020)) + 
    labs(title = "Overall carbon emissions by year by the major countries",
         subtitle = "Chinas emissions have taken off since the new millenium",
         x = "Year",
         y = "Kilotons of Carbon Dioxide Emitted",
         caption = "Source: UNFCC",
         color = "Country") + 
    theme_minimal() + 
    theme(
    legend.position = "right",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    axis.ticks.x = element_line(color = "black"),
    aspect.ratio = 0.6
  )
```

Russia used to be known as the USSR up until 1990 when it disbanded into Russia and several other countries. This took a toll on Russia's income and industrialization, resulting in less Carbon Emissions as shown on the graph. Additionally, the United States has been on somewhat a downward slope since about 2005. This is due to the emergence of renewable-energy and partly due to the Stock Market Crash of 2008, which limited the economy resulting in less production of material goods. China's emissions have increased the most in the last 20 years and it seems they will continue to increase in the years coming.