---
title: "Carbon Dioxide Emissions by Country (2019)"
author: "Tanay Janmanchi"
---

```{r}
#| label: setup
#| message: FALSE
#| echo: FALSE
library(tidyverse)
library(leaflet)
library(ggplot2)
library(sf)
library(dplyr)
library(viridis)
library(maps)
library(countrycode)
library(mapproj)
```

```{r}
#| label: data-download
#| message: FALSE
#| echo: FALSE

emissions <- read_csv(file = "carbon-emissions.csv", show_col_types = FALSE)
emissions <- emissions |> filter(year == 2019,
                                 country_code != "WLD",
                                 country_code != "IBT",
                                 country_code != "LMY",
                                 country_code != "MIC",
                                 country_code != "IBD",
                                 country_code != "UMC",
                                 country_code != "LTE",
                                 country_code != "EAS",
                                 country_code != "EAP",
                                 country_code != "TEA",
                                 country_code != "HIC",
                                 country_code != "OED",
                                 country_code != "PST",
                                 country_code != "EAR",
                                 country_code != "ECS",
                                 country_code != "LMC",
                                 country_code != "WLD",
                                 country_code != "TEC",
                                 country_code != "ECA",
                                 country_code != "SAS",
                                 country_code != "NAC",
                                 country_code != "TSA",
                                 country_code != "EUU",
                                 country_code != "MEA",
                                 country_code != "EMU",
                                 country_code != "ARB",
                                 country_code != "LCN",
                                 country_code != "TLA",
                                 country_code != "MNA",
                                 country_code != "TMN",
                                 country_code != "LAC",
                                 country_code != "IDA",
                                 country_code != "SSF",
                                 country_code != "TSS",
                                 country_code != "FCS",
                                 country_code != "AFE",
                                 country_code != "AFW",
                                 country_code != "CEB",
                                 country_code != "CSS",
                                 country_code != "IDB",
                                 country_code != "LDC",
                                 country_code != "LIC",
                                 country_code != "OSS",
                                 country_code != "PSS",
                                 country_code != "SST",
                                 country_code != "HBC",
                                 country_code != "PRE",
                                 country_code != "IDX",
                                 country_code != "SSA",
                                 country_code != "HPC")

emissions$country_name <- countrycode(emissions$country_code, origin = "iso3c", destination = "country.name")
emissions$country_name[emissions$country_code == "USA"] <- "USA"
emissions$country_name[emissions$country_code == "GBR"] <- "UK"
emissions$country_name[emissions$country_code == "DOM"] <- "DRC"
world_map <- map_data("world")
co2_map_data <- left_join(world_map, emissions, by = c("region" = "country_name"))


```

```{r}
#| label: plot-heatmap
#| message: false
#| echo: false
#| warning: false
#| fig-width: 15
#| fig-height: 12

co2_map_data |> ggplot(aes(x = long, y = lat, group = group, fill = value)) +
  geom_polygon(color = "black", linewidth = 0.2) +
  scale_fill_viridis(name = "Emissions (KtCO2e)", 
                     label = scales::comma, 
                     option = "viridis", 
                     direction = -1) +
  labs(
       subtitle = "Grey means data not available",
       x = NULL,
       y = NULL) + coord_map("mollweide", xlim = c(-180, 180), ylim = c(-90, 90)) + 
  theme_void() +
  theme(plot.subtitle = element_text(hjust = 0.5)) 

```

How does each country contribute to the Carbon Emissions of the world? The data for this graph was sourced from the United Nations Framework Convention on Climate Change (UNFCCC) and the International Energy Agency (IEA). After that, the data was visualized into a heatmap which shows emissions by country based on the color that is filled in. We can see that China (10,707,220 ktco2e) and the USA (4,817,720 ktco2e) emit the most Carbon Dioxide into our atmosphere, along with other countries in the Eastern Hemisphere known for high levels of industrialization and pollution, such as India, Russia and Indonesia.
