---
title: "Carbon Dioxide Emissions by Country (2019)"
author: "Tanay Janmanchi"
---

In the year 2019, 34,344,006 ktco2e were emitted into the atmosphere, do you wonder how each country contributed to the Carbon Emissions? We can see that China (10,707,220 ktco2e) and the USA (4,817,720 ktco2e) emit the most Carbon Dioxide into our atmosphere, along with other countries in the Eastern Hemisphere known for high levels of industrialization and pollution, such as India, Russia and Indonesia. Therefore, we can come to the conclusion that Carbon emissions can be credited to industrialization.

```{r}
#| label: setup
#| message: FALSE
#| echo: FALSE
# loading in the libraries
library(tidyverse)
library(leaflet)
library(ggplot2)
library(sf)
library(dplyr)
library(viridis)
library(maps)
library(countrycode)
library(mapproj)
```

```{r}
#| label: data-download
#| message: FALSE
#| echo: FALSE
# reading in the data
emissions <- read_csv(file = "data/carbon-emissions.csv", 
                      show_col_types = FALSE)
# filtering out any bad/unuseful data
emissions <- emissions |> filter(year == 2019,
                                 country_code != "WLD",
                                 country_code != "IBT",
                                 country_code != "LMY",
                                 country_code != "MIC",
                                 country_code != "IBD",
                                 country_code != "UMC",
                                 country_code != "LTE",
                                 country_code != "EAS",
                                 country_code != "EAP",
                                 country_code != "TEA",
                                 country_code != "HIC",
                                 country_code != "OED",
                                 country_code != "PST",
                                 country_code != "EAR",
                                 country_code != "ECS",
                                 country_code != "LMC",
                                 country_code != "WLD",
                                 country_code != "TEC",
                                 country_code != "ECA",
                                 country_code != "SAS",
                                 country_code != "NAC",
                                 country_code != "TSA",
                                 country_code != "EUU",
                                 country_code != "MEA",
                                 country_code != "EMU",
                                 country_code != "ARB",
                                 country_code != "LCN",
                                 country_code != "TLA",
                                 country_code != "MNA",
                                 country_code != "TMN",
                                 country_code != "LAC",
                                 country_code != "IDA",
                                 country_code != "SSF",
                                 country_code != "TSS",
                                 country_code != "FCS",
                                 country_code != "AFE",
                                 country_code != "AFW",
                                 country_code != "CEB",
                                 country_code != "CSS",
                                 country_code != "IDB",
                                 country_code != "LDC",
                                 country_code != "LIC",
                                 country_code != "OSS",
                                 country_code != "PSS",
                                 country_code != "SST",
                                 country_code != "HBC",
                                 country_code != "PRE",
                                 country_code != "IDX",
                                 country_code != "SSA",
                                 country_code != "HPC")

#importing in the map
emissions$country_name <- countrycode(emissions$country_code, origin = "iso3c", destination = "country.name")

#manually changing some of the country codes
emissions$country_name[emissions$country_code == "USA"] <- "USA"
emissions$country_name[emissions$country_code == "GBR"] <- "UK"
emissions$country_name[emissions$country_code == "DRC"] <- "DOM"

#joining the map with the current data set
world_map <- map_data("world")
co2_map_data <- left_join(world_map, emissions, by = c("region" = "country_name"))


```

```{r}
#| label: plot-heatmap
#| message: false
#| echo: false
#| warning: false

# dimensions for graph
#| fig-width: 20
#| fig-height: 16

#plot
co2_map_data |> ggplot(aes(x = long, y = lat, group = group, fill = value)) +
  geom_polygon(color = "black", linewidth = 0.2) +
  scale_fill_viridis(name = "Emissions (KtCO2e)", 
                     label = scales::comma, 
                     option = "viridis", 
                     direction = -1) +
  labs(
       subtitle = "Grey means data not available",
       x = NULL,
       y = NULL, 
       caption = "Source: UNFCC") + 
  coord_map("mollweide", xlim = c(-180, 180), ylim = c(-90, 90)) + 
  theme_void() +
  theme(plot.subtitle = element_text(hjust = 0.5)) 

```


